### 2022-06-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@b8ef26e57717e6fa9014a0df28be203a8e88aca9 🚀

In fig.~\ref{fig:biasco}, we present the response of two thresholding
measures (PR-AUC and ROC-AUC) and two ranking measures (Informedness and
MCC) to a grid of 35 values of training set balance, and 35 values of
connectance, for the four component models as well as the ensemble.
ROC-AUC is always high, and does not vary with training set balance. On
the other hand, PR-AUC shows very strong responses, increasing with
training set balance. It is notable here that two classifiers that
seemed to be performing well (Decision Tree and Random Forest) based on
their MCC are not able to reach a high PR-AUC even at higher
connectances.All models reached a higher performance on more connected
networks, and using more balanced training sets. In all cases,
informedness was extremely high, which is an expected consequence of the
fact that this is the value we optimized to determine the cutoff. MCC
increased with training set balance, although this increase became less
steep with increasing connectance. Three of the models (kNN, decision
tree, and random forest) only increased their PR-AUC sharply when the
training set was heavily imbalanced towards more interactions.
Interestingly, the ensemble almost always outclassed its component
models. For larger connectances (less difficult networks to predict, as
they are more balanced), MCC and informedness stared decreasing when the
training set bias got too close to one, suggesting that a training set
balance of 0.5 may often be appropriate if these measures are the one to
optimize.

@@

When trained at their optimal training set balance, connectance still
had a significant impact on the performance of some machines
fig.~\ref{fig:optimvalue}. Notably, Decision Tree, and k-NN, as well as
Random forest to a lower extent, had low values of PR-AUC. In all cases,
the Boosted Regression Tree was reaching very good predictions
(especially for connectances larger than 0.1), and the ensemble was
almost always scoring perfectly. This suggests that all the models are
biased in different ways, and that the averaging in the ensemble is able
to correct these biases. We do not expect this last result to have any
generality, and provide a discussion of a recent example in which the
ensemble was performing worse than its components models.

@@

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness
(\(\eta\); \textbf{Bastolla2009ArcMut?}), modularity (\(Q\);
\textbf{Barber2007ModCom?}), asymmetry (\(A\);
\textbf{Delmas2018AnaEco?}), and Jaccard network dissimilarity (Canard
et al., 2014). This process was repeated 250 times, and the results are
presented in tbl.~\ref{tbl:comparison}. The k-NN model is an interesting
instance here: it produces the network that looks the most like the
original dataset, despite having the lowest PR-AUC, suggesting it hits
high recall at the cost of low precision. The ensemble was able to reach
a very high PR-AUC (and a very high ROC-AUC), which translated into more
accurate reconstructions of the structure of the network (with the
exception of modulairty, which is underestimated by \(0.03\)). This
result bears elaborating. Measures of model performance capture how much
of the interactions and non-interactions are correctly identified. As
long as these predictions are not perfect, some interactions will be
predicted at the ``wrong'' position in the network; these measures
cannot describe the structural effect of these mistakes. On the other
hand, measures of network structure can have the same value with
interactions that fall at drastically different positions; this is in
part because a lot of these measures covary with connectance, and in
part because as long as these values are not 0 or their respective
maximum, there is a large number of network configurations that can have
the same value. That ROC-AUC is consistently larger than PR-AUC may be a
case of this measure masking models that are not, individually, strong
predictors (Jeni et al., 2013). In this specif example, the combination
of individually ``adequate'' models resulted in an extremely strong
ensemble, suggesting that the correct prediction of interactions (as
measured by MCC, Inf., ROC-AUC, and PR-AUC) \emph{and} network
properties is indeed a feasible task under appropriately
hyper-parameterized models.

@@

The results presented here highlight an interesting paradox: although
the k-NN model was ultimately able to get a correct estimate of network
structure tbl.~\ref{tbl:comparison}, it ultimately remains a poor
classifier, as evidenced by its low PR-AUC. This suggests that the goal
of predicting \emph{interactions} and predicting \emph{networks} may not
always be solvable in the same way -- of course a perfect classifier of
interactions would make a perfect network prediction; indeed, the best
scoring predictor of interactions (the ensemble model) had the best
prediction of network structure. The tasks of predicting networks
structure and of predicting interactions within networks are essentially
two different ones. For some applications (\emph{.e.g.} comparison of
network structure across gradients), one may care more about a robust
estimate of the structure, at the cost at putting some interactions at
the wrong place. For other applications (\emph{e.g.} identifying pairs
of interacting species), one may conversely care more about getting as
many pairs right, even though the mistakes accumulate in the form of a
slightly worse estimate of network structure. How these two approaches
can be reconciled is something to evaluate on a case-by-case basis,
especially since there is no guarantee that an esemble model will always
be the most precise one. Despite this apparent tension at the heart of
the predictive exercise, we can use the results presented here to
suggest a number of guidelines.

First, because we have more trust in reported interactions than in
reported absences of interactions (which are overwhelmingly
\emph{pseudo}-absences), we can draw on previous literature to recommend
informedness as a measure to decide on a threshold for binary
classification (Chicco et al., 2021); this being said, because
informedness is insensitive to bias (although it is a linear measure of
skill), the overall model performance is better evaluated through the
use of MCC fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously
sensitive to classifier bias fig.~\ref{fig:bias} and network connectance
fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
evaluation and comparison. When dealing with multiple models, we
therefore suggest to find the optimal threshold using informedness, and
to pick the best model using MCC (assuming one does not want to use an
ensemble model).

@@

Third, because the PR-AUC responds more to network connectance
fig.~\ref{fig:optimvalue} and training set imbalance
fig.~\ref{fig:biasrocpr} than ROC-AUC, it should be used as a measure of
model performance over the ROC-AUC. This is not to say that ROC-AUC
should be discarded (in fact, a low ROC-AUC is undoubtedly a sign of an
issue with the model), but that its interpretation should be guided by
the PR-AUC value. Specifically, a high ROC-AUC is not informative, as it
can be associated to a low PR-AUC (see \emph{e.g.} Random Forest in
tbl.~\ref{tbl:comparison}) This again echoes recommendations from other
fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015). We therefore
expect to see high ROC-AUC values, and then to pick the model that
maximizes the PR-AUC value. Taken together with the previous two
guidelines, we strongly encourage to (i) ensure that accuracy and
ROC-AUC are high (in the case of accuracy, higher than expected under
no-skill no-bias situation), and (ii) to discuss the performance of the
model in terms of the most discriminant measures, \emph{i.e.} PR-AUC and
MCC.

Finally, network connectance (\emph{i.e.} the empirical class imbalance)
should inform the composition of the training and testing set, because
it is an ecologically relevant value. In the approach outlined here, we
treat the class imbalance of the training set as an hyper-parameter, but
\emph{test} the model on a set that has the same class imbalance as the
actual dataset. This is an important distinction, as it ensure that the
prediction environment matches the testing environment (as we cannot
manipulate the connectance of the empirical dataset on which the
predictions will be made), and so the values measured on the testing set
(or validation set if the data volume allows one to exists) can be
directly compared to the values for the actual prediction. A striking
result fig.~\ref{fig:optimbias} is that Informedness was almost always
maximal at 50/50 balance (regardless of connectance), whereas MCC
required \emph{more} positives to be maximized when connectance
\emph{increases}, matching the idea that it is a more stringent measure
of performance. This has an important consequence in ecological
networks, for which the pool of positive cases (interactions) to draw
from is typically small: the most parsimonious measure (\emph{i.e.} the
one requiring to discard the least amount of interactions to train the
model) will give the best validation potential, and in this light is
very likely informedness (maximizing informedness is, in fact, the
generally accepted default for imbalanced classification regardless of
the problem domain; Schisterman et al., 2005). This last result further
strengthens the assumption that the amount of bias \emph{is} an
hyper-parameter that must be fine-tuned, as using the wrong bias can
lead to models with lower performance; for this reason, it makes sense
to not train all models on the same training/testing set, but rather to
optimize the set composition for each of them.

@@

@@

>>>>

In fig.~\ref{fig:biasco}, we present the response of two thresholding
measures (PR-AUC and ROC-AUC) and two ranking measures (Informedness and
MCC) to a grid of 35 values of training set balance, and 35 values of
connectance, for the four component models as well as the ensemble.
ROC-AUC is always high, and does not vary with training set balance. On
the other hand, PR-AUC shows very strong responses, increasing with
training set balance. It is notable here that two classifiers that
seemed to be performing well (Decision Tree and Random Forest) based on
their MCC are not able to reach a high PR-AUC even at higher
connectances. All models reached a higher performance on more connected
networks, and using more balanced training sets. In all cases,
informedness was extremely high, which is an expected consequence of the
fact that this is the value we optimized to determine the cutoff. MCC
increased with training set balance, although this increase became less
steep with increasing connectance. Three of the models (kNN, decision
tree, and random forest) only increased their PR-AUC sharply when the
training set was heavily imbalanced towards more interactions.
Interestingly, the ensemble almost always outclassed its component
models. For larger connectances (less difficult networks to predict, as
they are more balanced), MCC and informedness stared decreasing when the
training set bias got too close to one, suggesting that a training set
balance of 0.5 may often be appropriate if these measures are the one to
optimize.

@@

When trained at their optimal training set balance, connectance still
had a significant impact on the performance of some machines
(fig.~\ref{fig:optimvalue}). Notably, Decision Tree, and k-NN, as well
as Random forest to a lower extent, had low values of PR-AUC. In all
cases, the Boosted Regression Tree was reaching very good predictions
(especially for connectances larger than 0.1), and the ensemble was
almost always scoring perfectly. This suggests that all the models are
biased in different ways, and that the averaging in the ensemble is able
to correct these biases. We do not expect this last result to have any
generality, and provide a discussion of a recent example in which the
ensemble was performing worse than its components models.

@@

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness
(\(\eta\); Bastolla et al., 2009), modularity (\(Q\); Barber, 2007),
asymmetry (\(A\); Delmas et al., 2018), and Jaccard network
dissimilarity (Canard et al., 2014). This process was repeated 250
times, and the results are presented in tbl.~\ref{tbl:comparison}. The
k-NN model is an interesting instance here: it produces the network that
looks the most like the original dataset, despite having the lowest
PR-AUC, suggesting it hits high recall at the cost of low precision. The
ensemble was able to reach a very high PR-AUC (and a very high ROC-AUC),
which translated into more accurate reconstructions of the structure of
the network (with the exception of modulairty, which is underestimated
by \(0.03\)). This result bears elaborating. Measures of model
performance capture how much of the interactions and non-interactions
are correctly identified. As long as these predictions are not perfect,
some interactions will be predicted at the ``wrong'' position in the
network; these measures cannot describe the structural effect of these
mistakes. On the other hand, measures of network structure can have the
same value with interactions that fall at drastically different
positions; this is in part because a lot of these measures covary with
connectance, and in part because as long as these values are not 0 or
their respective maximum, there is a large number of network
configurations that can have the same value. That ROC-AUC is
consistently larger than PR-AUC may be a case of this measure masking
models that are not, individually, strong predictors (Jeni et al.,
2013). In this specif example, the combination of individually
``adequate'' models resulted in an extremely strong ensemble, suggesting
that the correct prediction of interactions (as measured by MCC, Inf.,
ROC-AUC, and PR-AUC) \emph{and} network properties is indeed a feasible
task under appropriately hyper-parameterized models.

@@

The results presented here highlight an interesting paradox: although
the k-NN model was ultimately able to get a correct estimate of network
structure (see tbl.~\ref{tbl:comparison} and fig.~\ref{fig:ecovalid}),
it ultimately remains a poor classifier, as evidenced by its low PR-AUC.
This suggests that the goal of predicting \emph{interactions} and
predicting \emph{networks} may not always be solvable in the same way --
of course a perfect classifier of interactions would make a perfect
network prediction; indeed, the best scoring predictor of interactions
(the ensemble model) had the best prediction of network structure. The
tasks of predicting networks structure and of predicting interactions
within networks are essentially two different ones. For some
applications (\emph{.e.g.} comparison of network structure across
gradients), one may care more about a robust estimate of the structure,
at the cost at putting some interactions at the wrong place. For other
applications (\emph{e.g.} identifying pairs of interacting species), one
may conversely care more about getting as many pairs right, even though
the mistakes accumulate in the form of a slightly worse estimate of
network structure. How these two approaches can be reconciled is
something to evaluate on a case-by-case basis, especially since there is
no guarantee that an esemble model will always be the most precise one.
Despite this apparent tension at the heart of the predictive exercise,
we can use the results presented here to suggest a number of guidelines.

First, because we have more trust in reported interactions than in
reported absences of interactions (which are overwhelmingly
\emph{pseudo}-absences), we can draw on previous literature to recommend
informedness as a measure to decide on a threshold for binary
classification (Chicco et al., 2021); this being said, because
informedness is insensitive to bias (although it is a linear measure of
skill), the overall model performance is better evaluated through the
use of MCC (figs.~\ref{fig:optimbias}, \ref{fig:optimvalue}). Because
\(F_1\) is monotonously sensitive to classifier bias fig.~\ref{fig:bias}
and network connectance fig.~\ref{fig:connectance}, MCC should be
prefered as a measure of model evaluation and comparison. When dealing
with multiple models, we therefore suggest to find the optimal threshold
using informedness, and to pick the best model using MCC (assuming one
does not want to use an ensemble model).

@@

Third, because the PR-AUC responds more to network connectance
(fig.~\ref{fig:optimvalue}) and training set imbalance
(fig.~\ref{fig:optimbias}) than ROC-AUC, it should be used as a measure
of model performance over the ROC-AUC. This is not to say that ROC-AUC
should be discarded (in fact, a low ROC-AUC is undoubtedly a sign of an
issue with the model), but that its interpretation should be guided by
the PR-AUC value. Specifically, a high ROC-AUC is not informative, as it
can be associated to a low PR-AUC (see \emph{e.g.} Random Forest in
tbl.~\ref{tbl:comparison}) This again echoes recommendations from other
fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015). We therefore
expect to see high ROC-AUC values, and then to pick the model that
maximizes the PR-AUC value. Taken together with the previous two
guidelines, we strongly encourage to (i) ensure that accuracy and
ROC-AUC are high (in the case of accuracy, higher than expected under
no-skill no-bias situation), and (ii) to discuss the performance of the
model in terms of the most discriminant measures, \emph{i.e.} PR-AUC and
MCC.

Finally, network connectance (\emph{i.e.} the empirical class imbalance)
should inform the composition of the training and testing set, because
it is an ecologically relevant value. In the approach outlined here, we
treat the class imbalance of the training set as an hyper-parameter, but
\emph{test} the model on a set that has the same class imbalance as the
actual dataset. This is an important distinction, as it ensure that the
prediction environment matches the testing environment (as we cannot
manipulate the connectance of the empirical dataset on which the
predictions will be made), and so the values measured on the testing set
(or validation set if the data volume allows one to exists) can be
directly compared to the values for the actual prediction. A striking
result from fig.~\ref{fig:optimbias} is that Informedness was almost
always maximal at 50/50 balance (regardless of connectance), whereas MCC
required \emph{more} positives to be maximized when connectance
\emph{increases}, matching the idea that it is a more stringent measure
of performance. This has an important consequence in ecological
networks, for which the pool of positive cases (interactions) to draw
from is typically small: the most parsimonious measure (\emph{i.e.} the
one requiring to discard the least amount of interactions to train the
model) will give the best validation potential, and in this light is
very likely informedness (maximizing informedness is, in fact, the
generally accepted default for imbalanced classification regardless of
the problem domain; Schisterman et al., 2005). This last result further
strengthens the assumption that the amount of bias \emph{is} an
hyper-parameter that must be fine-tuned, as using the wrong bias can
lead to models with lower performance; for this reason, it makes sense
to not train all models on the same training/testing set, but rather to
optimize the set composition for each of them.

@@

\leavevmode\hypertarget{ref-Barber2007ModCom}{}%
Barber, M. J. (2007). Modularity and community detection in bipartite
networks. \emph{Physical Review E}, \emph{76}(6), 066102.
\url{https://doi.org/10.1103/PhysRevE.76.066102}

\leavevmode\hypertarget{ref-Bastolla2009ArcMut}{}%
Bastolla, U., Fortuna, M. A., Pascual-García, A., Ferrera, A., Luque,
B., \& Bascompte, J. (2009). The architecture of mutualistic networks
minimizes competition and increases biodiversity. \emph{Nature},
\emph{458}(7241), 1018--1020. \url{https://doi.org/10.1038/nature07950}

@@

\leavevmode\hypertarget{ref-Delmas2018AnaEco}{}%
Delmas, E., Besson, M., Brice, M.-H., Burkle, L. A., Dalla Riva, G. V.,
Fortin, M.-J., Gravel, D., Guimarães, P. R., Hembry, D. H., Newman, E.
A., Olesen, J. M., Pires, M. M., Yeakel, J. D., \& Poisot, T. (2018).
Analysing ecological networks of species interactions. \emph{Biological
Reviews}, 112540. \url{https://doi.org/10.1111/brv.12433}
